Generalized spaces for constructive algebra

### 2020-04-08 paper-herrsching2019/main.tex
More references

\paragraph{Acknowledgments.} We are grateful to Peter Arndt, Andrej Bauer,
Martin Brandenburg, Thierry Coquand, Martín Escardó, Simon Henry, Matthias
Hutzler, Milly Maietti, Marc Nieper-Wißkirchen, Alexander Oldenziel, Anja
Petković, Peter Schuster, Helmut Schwichtenberg, Steve Vickers and Daniel
Wessel for invaluable discussions shaping this work. We thank the
organizers Klaus Mainzer, Peter Schuster and Helmut Schwich\-ten\-berg of
\emph{Proof and Computation 2019}, where this work was presented, for their
kind invitation and for creating a beautiful and stimulating meeting. We also
express our thanks to the local coordinator Chuangjie Xu and to all
participants of that meeting.

@@

For many purposes, the theory of prime \emph{filters} is actually more
relevant. Classically, a prime filter of a ring is simply the complement of a
prime ideal; constructively, it is prudent to turn prime ideals upside down and axiomatize this notion directly.
Hence the theory of prime filters of~$A$ has no sorts, no function symbols, one
relation symbol~``$D(x)$'' for each element~$x \in A$ and the following
axioms:
\begin{align*}
  D(0) &\vdash \bot \\
  D(x+y) &\vdash D(x) \vee D(y) & \text{(for each~$x,y \in A$)} \\
  D(xy) &\vdash D(x) & \text{(for each~$x,y \in A$)} \\
  \top &\vdash D(1) \\
  D(x) \wedge D(y) &\vdash D(xy) & \text{(for each~$x,y \in A$)}
\end{align*}

@@

We now begin the formal development of the sheaf semantics. More background and
examples from theoretical computer science, analysis and differential geometry
are contained in Ref.~\refcite{blechschmidt:custom-tailored}, which should be
regarded as a companion paper to this contribution.

@@

>>>>

@@

For many purposes, the theory of prime \emph{filters} is actually more
relevant.\footnote{This is not to say that all the issues with prime ideals in
constructive mathematics evaporate when turning to prime filters. For a time it
was believed that algebraic geometry could be developed constructively if only
the spectrum of a ring was defined as the topological space of its prime
filters~\cite[Section~3]{lawvere:icm-address}, but this hope turned out to be
too naive. André Joyal gave an explicit example of a nontrivial sheaf model of
the theory of rings without any prime filters~\cite[pp.~200f.]{tierney:spectrum}.}
Classically, a prime filter of a ring is simply the complement of a
prime ideal; constructively, it is prudent to turn prime ideals upside down and axiomatize this notion directly.
Hence the theory of prime filters of~$A$ has no sorts, no function symbols, one
relation symbol~``$D(x)$'' for each element~$x \in A$ and the following
axioms:
\begin{align*}
  D(0) &\vdash \bot \\
  D(x+y) &\vdash D(x) \vee D(y) & \text{(for each~$x,y \in A$)} \\
  D(xy) &\vdash D(x) & \text{(for each~$x,y \in A$)} \\
  \top &\vdash D(1) \\
  D(x) \wedge D(y) &\vdash D(xy) & \text{(for each~$x,y \in A$)}
\end{align*}

@@

We now begin the formal development of the sheaf semantics. More background and
examples from theoretical computer science, analysis and differential geometry
are contained in Ref.~\refcite{blechschmidt:custom-tailored}, which should be
regarded as a companion paper to this contribution. We also recommend
Christopher Mulvey's tutorial~\cite{mulvey:repr}, which culminates in a proof
of the Serre--Swan theorem using the sheaf semantics. Standard references for
the sheaf semantics include Refs.
\cite[Chapter~VI]{moerdijk-maclane:sheaves-logic},
\cite[Chapter~14]{goldblatt:topoi},
\cite[Section~5]{caramello:preliminaries},
\cite[Sections~12--15]{streicher:ctcl},
\cite[Chapter~6]{borceux:handbook3},
\cite[Part~D]{johnstone:elephant} and~[\refcite{shulman:categorical-logic}],
which also gives a detailed account of useful type theory.

@@

\section*{Acknowledgments}

We are grateful to Peter Arndt, Andrej Bauer,
Martin Brandenburg, Thierry Coquand, Martín Escardó, Simon Henry, Matthias
Hutzler, Milly Maietti, Marc Nieper-Wißkirchen, Alexander Oldenziel, Anja
Petković, Peter Schuster, Helmut Schwichtenberg, Steve Vickers and Daniel
Wessel for invaluable discussions shaping this work.

We thank the
organizers Klaus Mainzer, Peter Schuster and Helmut Schwich\-ten\-berg of
\emph{Proof and Computation 2019}, where this work was presented, for their
kind invitation and for creating a beautiful and stimulating meeting. We also
express our thanks to the local coordinator Chuangjie Xu and to all
participants of that meeting.
